---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, 
"SELECT * 
FROM order_products__prior_tbl 
LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl 
LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}
products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
dbGetQuery(sc, 
"
SELECT order_dow
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_dow
ORDER BY n_orders DESC
")
```

* The most popular hour of doing order? 
* The most reordered products?
* The less reorderd product?
* Plot histogram of number of products in an order.

```{r}
dbGetQuery(sc, 
"SELECT order_id
,   COUNT(1) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id
LIMIT 100
")
```

* What is the most popular department? 

```{r}
dbGetQuery(sc, "DESC orders_tbl")
```


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
```{r}
department_by_id_txt <- "select * from products_tbl where department_id = {{department_id}}
"

data <- list(department_id = "19")

department_by_id_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc,.)

```

* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}
"
SELECT pn.product_id, o.product_name,   
AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
    LEFT JOIN (
            SELECT  product_id, product_name
            FROM products_tbl) o
    ON o.product_id = pn.product_id
GROUP BY pn.product_id, o.product_name
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name? Did (left join)


Final project.
SQL sentences

1) Products more ordered by day of the week and it's quantity
```{r}
d <-"select a.*, p.product_name
from(
  select o.order_dow, opp.product_id, count(opp.product_id) quantity
  from order_products__prior_tbl  opp
  left join
    (select order_dow, order_id
    from orders_tbl
    ) o
  on opp.order_id=o.order_id
  group by order_dow, opp.product_id) a 
join 
  (select max(quantity) as max_quantity from 
    (select o.order_dow, opp.product_id, count(opp.product_id) quantity
      from order_products__prior_tbl  opp
      left join
        (select order_dow, order_id
        from orders_tbl
        ) o
      on opp.order_id=o.order_id
      group by order_dow, opp.product_id) a
  group by order_dow) b
on a.quantity = b.max_quantity
join
  (select product_name, product_id
  from products_tbl) p
  on a.product_id = p.product_id
order by order_dow" %>%
  dbGetQuery(sc,.)

d %>%
  ggplot(
    aes(reorder(order_dow, quantity, function(x) x), 
        quantity, fill=product_name)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("order_dow") +
  ylab("quantity") +
  theme_minimal()



```

2) 5 users with max number of products in a order
```{r}

d <- "select user_id, a.order_id, number_products from orders_tbl b
join
(select order_id, count(product_id) number_products from order_products__prior_tbl
  group by order_id) a
on b.order_id = a.order_id
order by number_products desc
limit 10" %>%
  dbGetQuery(sc,.)

d %>%
  ggplot(
    aes(reorder(user_id, number_products), 
        number_products, fill=user_id)) +
  geom_bar(stat="identity") + 
  xlab("User") +
  ylab("Number of products") +
  theme_minimal()
```

